\ttitle

### 2014-01-10 Chapters/Chapter1.tex
Finalizing Chapter 2, literatur review...

In this thesis the goal is to find the temporal segmentation of time series data explicitly, prior to the classification of the activities.
This is done under the assumption that it could be beneficial for classification methods to possess this explicit segmentation, since the model construction phase can use more data than the time-windowed approach would allow.
To find a temporal segmentation, we rely on change detection in time series data.
In the context of this research, this time series data consists of recordings from inertial sensors found in smartphones, such as the accelerometer signal, during the performance of human activities.
We are interested in both in- and outdoor activities in an uncontrolled environment and in a continuous manner, which can be regarded as real-world data.
Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust then in the case of data from a controlled environment.

For the change detection algorithm we have used a special form of \gls{svm}, the \gls{occ} based \gls{svdd} as introduced by Tax~\cite{tax2001one}.
This method models the data under consideration in the shape of a high-dimensional hypersphere.
From this hypersphere the radius is extracted, which can be used as an indication of change and be processed by one-dimensional change detection algorithms.
Since a change in data characteristics, resulting from a change in performed activity, alters the data distribution, the model and thus the radius should reflect this change as well.

\section{Scope}
The scope of this thesis is to find a temporal segmentation of time series data from inertial sensors embedded in smartphones, obtained during the performance of human activities.
As mentioned above, the segmentation is often obtained as a by-product from direct windowed classification.
Other approaches have a very rigid form of segmentation, \eg by limiting the number of possible segment types \cite{himberg2001time,chamroukhi2013joint}, or have applied it to different type of time series data \cite{li2007segmentation,fuchs2010online,guenterberg2009automatic}.
Others have extracted features from windows of data which are used for the segmentation \cite{guo2012adaptive}.
These activities include sitting, standing, walking, running, and ascending and descending staircases.
The problem of finding a temporal segmentation can be seen as a sub-problem to the problem of Activity Classification.
This relation is illustrated in \Cref{fig:thesis_goal}.
The figure shows that we can apply classification over windows of the time series data.
In many approaches this is constructed with a (partially overlapping) sliding window.
In the context of \gls{har}, the window length commonly consists of around $2$ seconds of data points.
In our setup, we create a temporal segmentation instead of a fixed window length.
Each segment should consist of homogeneous data, in the sense that the underlying data generating activity should not change within a segment.
Likewise, the previous and next segments (if present) should be generated by a different performed activity.

@@

\section{One-Class Support Vector Machines}
In the setup, as described in the previous section, we employ a model construction phase.
For this model we will use an \acrlong{oc-svm} implementation: the \acrlong{svdd} algorithm by Tax~\cite{tax2001one}.
In earlier research multi-class \glspl{svm} are used for direct classification of inertial sensor data \cite{he2008activity,mountrakis2011support,anguita2012human}.
Others have used \gls{oc-svm} classifiers for novelty, or outlier, detection \cite{scholkopf1999support,camci2010change,li2003improving,ma2003time,tax1999support}.
The research by Yin \etal  \cite{yin2008sensor} creates a \gls{oc-svm} classifier to filter normal activity traces, using training data.
However, to the best of our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly to find changes in behavior and create a temporal segmentation.
This algorithm transforms the time series data, which can be of any high dimension, to an higher dimensional feature space.
When the data is assumed to be of the same class, it can be characterized by an enclosing boundary, in the shape of a hypersphere.
Using this \gls{occ} method, we are able to test new data points of being a member of this class.
When a new data point, after transformation to the feature space, lies within the boundary, it is a member of the class, and otherwise it is regarded as an outlier.

@@

\section{Data sets}
In recent previous research two data sets for the context of \gls{har} have been made public.
In order to benchmark our proposed method we tried to employ it to these data sets.
However, it turned out the data sets were not appropriate for continuous temporal segmentation.
The first set, the \gls{wisdm} from \cite{kwapisz2011activity}, consists of non-continuous recordings of activities.
Obviously, this voids the usefulness of the data set of our research.
The second set, the \gls{uci-har} from \cite{anguita2012human}, also lacks continuous recordings and furthermore exhibits patterns and labeling which seems to be incorrect.
Besides these problems, since there are no visual images of the performed activities, there is no ground truth which can be consulted in case of ambiguity.
To overcome these problems, we have created our own real-world data sets.

>>>>

In this thesis the goal is to find the temporal segmentation of time series data explicitly, prior to the classification of the activities.
This is done under the assumption that it could be beneficial for classification methods to possess this explicit segmentation, since the model construction phase can use more data than the time-windowed approach would allow.
To find a temporal segmentation, we rely on change detection in time series data.
In the context of this research, this time series data consists of recordings from inertial sensors found in smartphones, such as the accelerometer signal, during the performance of human activities.
We are interested in both in- and outdoor activities in an uncontrolled environment and in a continuous manner, which can be regarded as real-world data.
Although the number of performed activities will be limited, we assume the temporal segmentation must be more robust than in the case of data from a controlled environment.

For the change detection algorithm we will use a special form of a \gls{svm}, the \gls{occ} based \gls{svdd}, as introduced by Tax~\cite{tax2001one}.
This method models the data under consideration in the shape of a high-dimensional hypersphere.
From this hypersphere the radius is extracted, which can be used as an indication of change and be processed by one-dimensional change detection algorithms.
Since a change in data characteristics, resulting from a change in performed activity, alters the data distribution, the model and thus the radius should reflect this change as well.
In the reverse relation, a change in the radius indicates a change in the data distribution and thus a change in the performed activity.

\section{Scope}
The scope of this thesis is to find a temporal segmentation of time series data from inertial sensors embedded in smartphones, obtained during the performance of human activities.
As mentioned above, the segmentation is often obtained as a by-product from direct windowed classification.
Other approaches have a very rigid form of segmentation, \eg by limiting the number of possible segment types \cite{himberg2001time,chamroukhi2013joint}, or have applied it to different type of time series data \cite{li2007segmentation,fuchs2010online,guenterberg2009automatic}.
Others have extracted features from windows of data which are used for the segmentation \cite{guo2012adaptive}.
These activities include sitting, standing, walking, running, and ascending and descending staircases.
The problem of finding a temporal segmentation can be seen as a sub-problem to the problem of Activity Classification.
This relation is illustrated in \Cref{fig:thesis_goal}.
The figure shows that we can apply classification over windows of the time series data.
In many approaches this is constructed with a (partially overlapping) sliding window.
In the context of \gls{har}, the window length commonly consists of around $2$ seconds of data points.
In our setup, we create a temporal segmentation instead of a fixed window length.
Each segment will consist of homogeneous data, in the sense that the underlying data generating activity should not change within a segment.
Likewise, the previous and next segments (if present) should be generated by a different performed activity.

@@

\section{One-Class Support Vector Machines}
In the setup, as described in the previous section and visualized in \Cref{fig:thesis_goal}, we employ a model construction phase.
For this model we will use an \acrlong{oc-svm} implementation: the \acrlong{svdd} algorithm by Tax~\cite{tax2001one}.
In earlier research multi-class \glspl{svm} are used for direct classification of inertial sensor data \cite{he2008activity,mountrakis2011support,anguita2012human}.
Others have used \gls{oc-svm} classifiers for novelty, or outlier, detection \cite{scholkopf1999support,camci2010change,li2003improving,ma2003time,tax1999support}.
The research by Yin \etal  \cite{yin2008sensor} creates a \gls{oc-svm} classifier to filter normal activity traces, using training data.
However, to the best of our knowledge, \gls{oc-svm} methods have not yet been applied to inertial sensor time series data directly to find changes in behavior and create a temporal segmentation.
This algorithm transforms the time series data, which can be of any high dimension, to an higher dimensional feature space.
When the data is assumed to be of the same class, it can be characterized by an enclosing boundary, in the shape of a hypersphere.
Using this \gls{occ} method, we are able to test new data points of being a member of this class.
When a new data point, after transformation to the feature space, lies within the boundary, it is a member of the class, and otherwise it is regarded as an outlier.

@@

\section{Data sets}
In recent previous research two data sets for the context of \gls{har} have been made public.
In order to benchmark our proposed method we tried to apply our method to these data sets.
However, it turned out the data sets were not appropriate for continuous temporal segmentation.
The first set, the \gls{wisdm} from \cite{kwapisz2011activity}, consists of non-continuous recordings of activities.
Obviously, this voids the usefulness of the data set of our research.
The second set, the \gls{uci-har} from \cite{anguita2012human}, also lacks continuous recordings and furthermore exhibits patterns and labeling which seems to be incorrect.
Besides these problems, since there are no visual images of the performed activities, there is no ground truth which can be consulted in case of ambiguity.
To overcome these problems, we have created our own real-world data sets.

### 2014-01-10 Chapters/Chapter2.tex
Finalizing Chapter 2, literatur review...

%----------------------------------------------------------------------------------------
\section{Outline}
\emph{Not intented for the reader.}
\begin{itemize}
  \item Literature review about Temporal Segmentation (previous draft was more about classification)
  \item Consider methods for the context of filter-methods for classification
  \item Take a loot at 3-4 different kind of methods for change detection:
    \begin{itemize}
      \item Introduction with a lot of techniques
      \item Explain why look at a few
      \item CUSUM - or other more traditional methods
      \item Density-ratio estimation
      \item Support Vector Machines (?) - if there are more sources about this
      \item (Dimensionality reduction) --> probably not
      \item \emph{Not to much attention to all techniques, focus is on SVM}
    \end{itemize}
  \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
\end{itemize}

\section{Statistical framework}\label{statistical-framework}

-- Nieuwe indeling Chapter 2 --

\begin{description}
  \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
  \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
  \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
  \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
    \begin{description}
      \item[Model fitting]
      \item[Cusum]
      \item[Density ratio]
    \end{description}
  \item[Support Vector Machine based methods]

\end{description}
\input{./Chapters/Chapter2/overview.tex}
\input{./Chapters/Chapter2/statistical_models.tex}
\input{./Chapters/Chapter2/temporal_segmentation.tex}
\input{./Chapters/Chapter2/change_detection_methods.tex}

-- OLD --

%----------------------------------------------------------------------------------------
% CUSUM and GLR
\input{./Chapters/Chapter2/cusum}

>>>>

%----------------------------------------------------------------------------------------
% \section{Outline}
% \emph{Not intented for the reader.}
% \begin{itemize}
%   \item Literature review about Temporal Segmentation (previous draft was more about classification)
%   \item Consider methods for the context of filter-methods for classification
%   \item Take a loot at 3-4 different kind of methods for change detection:
%     \begin{itemize}
%       \item Introduction with a lot of techniques
%       \item Explain why look at a few
%       \item CUSUM - or other more traditional methods
%       \item Density-ratio estimation
%       \item Support Vector Machines (?) - if there are more sources about this
%       \item (Dimensionality reduction) --> probably not
%       \item \emph{Not to much attention to all techniques, focus is on SVM}
%     \end{itemize}
%   \item With each method, shortly look at characteristics, strengths and weaknesses and consider applicability to accelerometer sensor data
% \end{itemize}

% \section{Statistical framework}\label{statistical-framework}

% -- Nieuwe indeling Chapter 2 --

% \begin{description}
%   \item[Intro/outline] Introduction and overview of this chapter. Put in the context of temporal classification.
%   \item[Temporal segmentation] Give overview of used temporal segmentation methods, move towards a split in model building/dimensionality reduction and change detection methods
%   \item[Dimensionality reduction] Explain that \gls{dr} can be useful in data processing. Name a few techniques, make bridge to \gls{oc-svm}.
%   \item[Change detection] Describe methods for \emph{direct} change detection. End with notion
%     \begin{description}
%       \item[Model fitting]
%       \item[Cusum]
%       \item[Density ratio]
%     \end{description}
%   \item[Support Vector Machine based methods] test

% \end{description}
As described in \Cref{Chapter1}, the focus of this research is the application of temporal segmentation to accelerometer data.
This is with the assumption that explicit temporal segmentation can aid the process of temporal classification.

\TODO{Give examples of temporal classification techniques}

In recent years a lot is research is performed in the classification of human activities, recorded by on-body sensors.
Especially since smartphones with accelerometer sensors became more available, research has focused on recordings from these devices.

\TODO{Include highlights from old-classification text}

In the following of this chapter we will start with an overview of statistical models used in this field of study.
That section will act as a starting point to look at the different types, applications and methods of temporal segmentation.
The next section, \Cref{sec:literature_review_temporal_segmentation} follows the distinction made and looks at a collection of temporal segmentation and change detection methods.
It will also relate the concept of \gls{occ} to temporal segmentation and change detection.
The final section of this chapter, \Cref{sec:literature_review_svm}, will discuss a specific class of algorithms for \gls{occ}, namely the methods that use \glspl{svm} for the classification task.

\input{./Chapters/Chapter2/statistical_models.tex}
\input{./Chapters/Chapter2/temporal_segmentation.tex}
% \input{./Chapters/Chapter2/change_detection_methods.tex}

% -- OLD --

%----------------------------------------------------------------------------------------
% CUSUM and GLR
% \input{./Chapters/Chapter2/cusum}

### 2014-01-10 Chapters/Chapter2/statistical_models.tex
Finalizing Chapter 2, literatur review...

% !TEX root = ../../main.tex
\section{Statistical models}\label{sec:literature_review_statistical_models}

\TODO{Introduce the diffent models and thus different approaches}

Many applications require the detection of time points at which the underlying properties of a system change.
This problem has received a lot of attention in the fields of data mining, etc... \TODO{list and refs}.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from benchmarks.
This method and some derivatiges are discussed and analyzed in \Cref{sec:cusum}.

@@

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci \cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hyper-sphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
\Cref{sec:literature_review_svm} discusses the \gls{svm}-methods.
\TODO{Put emphasis that this is the source of inspiration for the chosen method?}

-- NEW --

In the field of temporal segmentation of time series data we can distinguish different approaches and common frameworks in which they perform.
The problem is commonly formulated in the statistical framework and regarding the properties of the underlying data generating \gls{pdf} fo the time series data.
In that sense, a change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean or variance.

@@

% Piecewise approximation

-- Density

-- Outlier based

>>>>

% !TEX root = ../../main.tex
\section{Statistical models}\label{sec:literature_review_statistical_models}
Many applications require the detection of time points at which the underlying properties of a system change.
Often this problem is formulated in a statistical framework, by inspecting the underlying data generating \gls{pdf} of the time series data.
A change point is then defined as a significant change in the properties of the \gls{pdf}, such as the mean and variance.

The widely used \gls{cusum} method by Basseville \etal \cite{basseville1993detection} follows this approach.
It originates from control methods for detection from benchmarks.
This method, and some derivatives, are discussed and analyzed in \Cref{subsec:literature_review_temporal_segmentation_cusum}.

@@

The density-estimation methods rely on the log-likelihood ratio between \glspl{pdf}.
The method of Camci~\cite{camci2010change} follows an other approach within the statistical framework, by using a \gls{svm}.
One problem it tries to overcome is the (claimed) weakness of many methods to detect a decrease in variance.
The method represents the distribution over the data points as a hypersphere in a higher dimension using the kernel trick.
A change in the \gls{pdf} is represented by a change in the radius of this sphere.
In \Cref{sec:literature_review_svm} more applications of the \gls{svm}-based methods are discussed.
Since \gls{occ} and the \gls{oc-svm} method is central in our approach, \Cref{sec:one_class_classification,sec:one_class_svm} discuss these techniques in detail.

@@

### 2014-01-10 Chapters/Chapter2/svm.tex
Finalizing Chapter 2, literatur review...

Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a (supervised) model learning method.
An elaborate overview of applications is given in \cite{mountrakis2011support}.
In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
A number of proposals have been made using this method.
Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
In the case of novelty detection, they argue there is no need for density estimation of the distribution.
Simple algorithms estimate the density by considering how many data points fall in a region of interest.
The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.
It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data by separating it from the origin.
The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.

@@

The inner workings of \glspl{oc-svm} are discussed in detail in \Cref{sec:one_class_svm}.

\TODO{Add notion of `supervised' vs `unsupervised' learning. \gls{oc-svm} is unsupervised.}

>>>>

Many proposals in the field of remote sensing and \gls{har} make use of \glspl{svm} as a (supervised) model learning method.
An elaborate overview of applications is given in \cite{mountrakis2011support}.
In this section we review methods of change detection based on the applications of \glspl{svm} as model construction method.
A number of proposals have been made using this method.
Sch{\"o}lkopf \etal \cite{scholkopf1999support} applies Vapnik's principle, to never solve a problem which is more general than the one that one is actually interested in, to novelty detection.
In the case of novelty detection, they argue there is no need for density estimation of the distribution.
Simple algorithms estimate the density by considering how many data points fall in a region of interest.
The \gls{nu-svm} method instead starts with the number of data points that should be within the region and estimates a region with that desired property.
It builds on the method of Vapnik and Vladimir \cite{vapnik1963pattern}, which characterizes a set of data points by separating it from the origin.
The \gls{nu-svm} method adds the kernel method, allowing non-linear decision functions, and incorporates `softness' by the $\nu$-parameter.
Whereas the method in \cite{vapnik1963pattern} focuses on two-class problems, the \gls{nu-svm} method introduces the method the one-class problems.

@@

The following chapter discusses \gls{occ}, \gls{oc-svm}, and change detection through \gls{oc-svm} in detail.
It will show that the used implementation of \gls{oc-svm} is an example of the Type 3 semi-supervised learning methods, in the system of Hodge and Austin~\cite{hodge2004survey}.

### 2014-01-10 Chapters/Chapter2/temporal_segmentation.tex
Finalizing Chapter 2, literatur review...

\TODO{Is there any research of temporal segmentation on accelerometer/inertia signals? Note if there is/isnt}

Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.

\begin{itemize}
  \item Given overview of segmentation techniques, for times series data
  \item Use different ``point-of-views'', or terms
  \item ``Segmentation''
  \item ``Change detection''
  \item ``Novelty detection''
  \item Specific view on \glspl{svm}
\end{itemize}

This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
It takes a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
These will be the categorical terms for which we discuss the literature.

Aspects we will consider when discussing the literature are applicability to \gls{har}, parameter tuning and ... \TODO{What more aspects?}
% Finally we will discuss other applications of \glspl{svm} in the context of these terms.

\subsection{Segmentation}\label{subsec:segmentation}

@@

Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the former method.
Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ration is increased, resulting in a robuster method.

The method of Guenterberg \etal \cite{guenterberg2009automatic} extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
Using the \gls{swab} method the self-similarity of a one-dimensional signal is obtained.
The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
After the segmentation phase, the method of Bernecker \etal extracts other statistical features which are used in the classification phase.

The proposal of Guo \etal \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
By extending the \gls{swab} method, for every frame a feature set is selected, using an enhanced version of \gls{pca}.
The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
Whereas the before mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.

@@

The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function if time is modeled by calculating the posterior probability.
It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.

@@

Many methods obtain an implicit segmentation as a result of classification over a sliding window \TODO{add refs}.
The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.

@@

Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.

@@

``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\

@@

``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\

@@

\TODO{Link this subsection with previous and next}
% Papers:
% \begin{itemize}
%   \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
%   Implemented this algorithm.
%   \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
%   \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
% \end{itemize}

\emph{Notes:}
Non-Bayesian change detection algorithm (\ie no prior distribution believe available for the change time).
The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
Primary for detection of mean shift.
The \gls{mosum} of squares test for monitoring variance changes \cite{hsu2007mosum}.
Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}

An often used approach in the statistical framework of change detection is the \gls{cusum} as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements and has been applied to this problem \cite{basseville1993detection}.
It is a non-Bayesian method and thus makes no assumptions, in the form of prior belief distributions, for the change points.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the \gls{mosum} of squares by \cite{hsu2007mosum}.

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
Using the batch \gls{icss} algorithm they are able to find multiple change points, offline while post-processing the data.
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum of squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.

@@

\subsection{Change detection}\label{subsec:change_detection}
\TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
Although these techniques are closely related, there is a subtle difference.
In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.

The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
Using an iterative approach, all possible change points are considered.
The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.

\Cref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

\TODO{move CUSUM based techniques to this subsection}

>>>>

% \TODO{Is there any research of temporal segmentation on accelerometer/inertia signals? Note if there is/isnt}

% Give overview of used temporal segmentation methods. Problem is that it is not really well-defined.

% \begin{itemize}
%   \item Given overview of segmentation techniques, for times series data
%   \item Use different ``point-of-views'', or terms
%   \item ``Segmentation''
%   \item ``Change detection''
%   \item ``Novelty detection''
%   \item Specific view on \glspl{svm}
% \end{itemize}

This section gives an overview of the literature on temporal segmentation in the context of \gls{har}.
It provides a look on different implementations and methodologies.
A wide range of terms and subtle differences are used in the field, such as `segmentation', `change detection', `novelty detection' and `outlier detection'.
The following sections discuss these different methods.

\subsection{Segmentation based methods}\label{subsec:segmentation}

@@

Other methods have been proposed, such as an adaptive threshold based on the signal energy by Guenterberg \etal \cite{guenterberg2009automatic}, the adaptive \gls{cusum}-based test by Alippi \etal \cite{alippi2006adaptive} and the \gls{mosum} by Hsu \cite{hsu2007mosum} in order to eliminate this user-dependency.
The latter of these methods is able to process the accelerometer values directly, although better results are obtained when features of the signal are processed, as done in the first method \cite{guenterberg2009automatic}.
Here the signal energy, mean and standard deviation are used to segment activities and by adding all the axial time series together, the Signal-To-Noise ratio is increased, resulting in a robuster method.

The method of Guenterberg \etal \cite{guenterberg2009automatic} extracts features from the raw sensor signal to base the segmentation on other properties than the pure values.
The method of Bernecker \etal \cite{bernecker2012activity} uses other statistical properties, namely autocorrelation, to distinguish periodic from non-periodic segments.
Using the \gls{swab} method the self-similarity of a one-dimensional signal is obtained.
The authors claim that only a slight modification is needed to perform the method on multi-dimensional data.
After the segmentation phase, the method of Bernecker \etal \cite{bernecker2012activity} extracts other statistical features which are used in the classification phase.

The proposal of Guo \etal \cite{guo2012adaptive} dynamically determines which features should be used for the segmentation and simultaneously determines the best model to fit the segment.
For each of the three dimensions features such as the mean, variance, covariance, correlation, energy and entropy are calculated.
By extending the \gls{swab} method, for every frame a feature set is selected, using an enhanced version of \gls{pca}.
The research also considered the (Stepwise) Feasable Space Window as introduced by \cite{liu2008novel}, but since it results in a higher error rate than \gls{swab} the latter was chosen to extend.
Whereas the above mentioned algorithms use a linear representation, this methods considers linear, quadratic and cubical representations for each segment.
This differs from other methods where the model is fixed for the whole time series, such as \cite{fuchs2010online}, which is stated to perform inferior on non-stationary time series such as daily life.

@@

The method of Adams and MacKay \cite{adams2007bayesian} builds a probabilistic model on the segment run length, given the observed data so far.
Instead of modeling the values of the data points as a probabilistic distribution, the length of segments as a function of time is modeled by calculating the posterior probability.
It uses a prior estimate for the run length and a predictive distribution for newly-observed data, given the data since the last change point.
This method contrasts with the approach of Guralnik and Srivastava \cite{guralnik1999event} in which change points are detected by a change in the (parameters of an) underlying, observed, model.
For each new data point, the likelihoods of being a change point and part of the current segment are calculated, without a prior model (and thus is a non-Bayesian approach).
It is observed that when no change point is detected for a long period of time, the computational complexity increases significantly.

@@

As stated before, oten methods obtain an implicit segmentation as a result of classification over a sliding window.
The method of Yang \etal \cite{yang2008distributed} explicitly performs segmentation and classification simultaneously.
It argues that the classification of a pre-segmented test-sequences becomes straightforward with many classical algorithms to choose from.
The algorithm matches test examples with the \emph{sparsest} linear representation of mixture subspace models of training examples, searching over different temporal resolutions.

@@

% Field of computer vision: \cite{zhou2008aligned}, \cite{li2007segmentation}.

@@

% ``Segmentation and Recognition of Motion Streams by Similarity Search'' \cite{li2007segmentation}. 29, 2007 \\

@@

% ``Aligned Cluster Analysis for Temporal Segmentation of Human Motion'' \cite{zhou2008aligned}. 63, 2008 \\

@@

% \TODO{Link this subsection with previous and next}
% Papers:
% \begin{itemize}
%   \item Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance, 1994, 162 refs \cite{inclan1994use} \\
%   Implemented this algorithm.
%   \item An adaptive CUSUM-based test for signal change detection, 2006, 15 refs. \cite{alippi2006adaptive}
%   \item The MOSUM of squares test for monitoring variance changes \cite{hsu2007mosum}
% \end{itemize}

% \emph{Notes:}
% Non-Bayesian change detection algorithm (\ie no prior distribution believe available for the change time).
% The \gls{cusum} method is developed by Page \cite{page1954continuous} for the application of statistical quality control (it is also known as a control chart).
% Primary for detection of mean shift.
% The \gls{mosum} of squares test for monitoring variance changes \cite{hsu2007mosum}.
% Use of Cumulative Sums of Squares for Retrospective Detection of Changes of Variance \cite{inclan1994use}

An other often used approach in the statistical framework of change detection is the \gls{cusum} as introduced by Page \cite{page1954continuous}.
Originally used for quality control in production environments, its main function is to detect change in the mean of measurements \cite{basseville1993detection}.
It is a non-Bayesian method and thus makes no assumptions, in the form of prior belief distributions, for the change points.
Many extensions to this method have been proposed.
Some focus on the change in mean, such the method of Alippi and Roveri \cite{alippi2006adaptive}.
Others apply the method the problems in which the change of variance is under consideration.
Among others are there the centered version of the cumulative sums, introduced by Brown, Durbin and Evans \cite{brown1975techniques} and the \gls{mosum} of squares by \cite{hsu2007mosum}.

The method of Incl\'{a}n and Tiao \cite{inclan1994use} builds on the centered version of \gls{cusum} \cite{brown1975techniques} to detect changes in variance.
The obtained results (when applied to stock data) are comparable with the application of \gls{mle}.
Using the batch \gls{icss} algorithm they are able to find multiple change points, offline while post-processing the data.
Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
Let $C_k = \sum_{i=1}^k \alpha_t^2$ be the cumulative sum of squares for a series of uncorrelated random variables $\{\alpha_t\}$ of length $T$.
The centered (and normalized) sum of squares is defined as
\begin{equation}
  \begin{aligned}
  D_k = \frac{C_k}{C_T} - \frac{k}{T}, & & k = 1, \dots, T, & & \text{with } D_0 = D_T = 0.
  \end{aligned}
\end{equation}
For a series with homogeneous variance, the value of $D_k$ will oscillate around $0$.
This $D_k$ is used to obtain a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
In case of a sudden change, the value will increase and exceed some predefined boundary with high probability.
By using an iterative algorithm, the method is able to minimize the masking effect of successive change points.
The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.

@@

% \subsection{Change detection methods}\label{subsec:change_detection}
% \TODO{change order of this and temporal segmentation sections? So first change, then segmentation?} \\
% Whereas the above mentioned researches focus on \emph{segmentation}, many have focused on \emph{change detection}.
% Although these techniques are closely related, there is a subtle difference.
% In the case of \emph{change detection} to goal is to find, possibly unrelated, sudden change points in a signal \cite{takeuchi2006unifying}.
% In contrast, the goal of \emph{temporal segmentation} is to find homogeneous segments of data, which can be the result of multiple detected changes.

% The \gls{icss} by Incl\'{a}n and Tiao \cite{inclan1994use} is a statistical method which obtains results (when applied to stock data) comparable to \gls{mle} and Bayesian \TODO{Bayesian What?}.
% Whereas \gls{cusum} can be applied to search for a change in mean, the \gls{icss} is adapted to find changes in variance.
% It obtains a \emph{likelihood ratio} for testing the hypothesis of one change against no change in the variance.
% Using an iterative approach, all possible change points are considered.
% The proposal of \cite{cheng2009efficient} extends on the \gls{cusum}-based methods to find change points in mean and variance, by creating a more efficient and accurate algorithm.

% \Cref{sec:change_detection_time_series} discusses the problem of change detection in time series further, and gives a formal problem definition.

% \TODO{move CUSUM based techniques to this subsection}
